---
title: "Working with Different Object Types and Themes"
author: "Ronald G. Thomas"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with Different Object Types and Themes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)

has_latex <- nzchar(Sys.which("pdflatex")) && nzchar(Sys.which("pdfcrop"))

output_dir <- tempdir()

include_table <- function(filename) {
  pdf_path <- file.path(output_dir, paste0(filename, "_cropped.pdf"))
  if (!file.exists(pdf_path)) return(invisible(NULL))
  png_path <- sub("\\.pdf$", ".png", pdf_path)
  convert_pdf_to_png(pdf_path, png_path, dpi = 150)
  knitr::include_graphics(png_path)
}
```

zzobj2fig v0.2.0 introduced S3 method dispatch for `o2f()`, enabling direct
conversion of statistical objects to publication-quality tables. This vignette
demonstrates how different R object types are handled and how themes affect
table appearance.

```{r load-package, message=FALSE, warning=FALSE}
library(zzobj2fig)
```


# Object Types

## Data Frames

Data frames are the most common input type. The `o2f.data.frame()` method
handles standard data frame objects with full formatting support.

```{r dataframe-example, eval=has_latex}
df <- data.frame(
  Variable = c("Age", "Weight", "Height"),
  Mean = c(45.2, 72.5, 168.3),
  SD = c(12.1, 15.8, 9.2),
  N = c(150, 148, 150)
)

o2f(df, filename = "descriptive_stats", sub_dir = output_dir,
    caption = "Descriptive Statistics")
```

```{r dataframe-display, echo=FALSE, eval=has_latex, out.width="50%"}
include_table("descriptive_stats")
```

Data frames from dplyr pipelines work seamlessly:

```{r dplyr-example, eval=has_latex, message=FALSE}
library(dplyr)

summary_table <- mtcars |>
  group_by(cyl) |>
  summarise(
    N = n(),
    `Mean MPG` = round(mean(mpg), 1),
    `SD MPG` = round(sd(mpg), 1),
    `Mean HP` = round(mean(hp), 0),
    .groups = "drop"
  )

o2f(summary_table,
    filename = "cylinder_summary",
    caption = "Vehicle Statistics by Cylinder Count",
    sub_dir = output_dir)
```

```{r dplyr-display, echo=FALSE, eval=has_latex, out.width="55%"}
include_table("cylinder_summary")
```


## Matrices

Matrices are converted to data frames with optional row names as the first
column via the `o2f.matrix()` method.

```{r matrix-example, eval=has_latex}
cor_mat <- cor(mtcars[, c("mpg", "cyl", "disp", "hp", "wt")])
cor_mat <- round(cor_mat, 2)

o2f(cor_mat,
    filename = "correlation_matrix",
    caption = "Correlation Matrix of Vehicle Characteristics",
    rownames = TRUE,
    sub_dir = output_dir)
```

```{r matrix-display, echo=FALSE, eval=has_latex, out.width="65%"}
include_table("correlation_matrix")
```


## Contingency Tables

Base R `table` objects are supported directly via the `o2f.table()` method.

```{r table-example, eval=has_latex}
gear_cyl_table <- table(mtcars$gear, mtcars$cyl)

o2f(gear_cyl_table,
    filename = "gear_cylinder_crosstab",
    caption = "Cross-tabulation of Gears and Cylinders",
    sub_dir = output_dir)
```

```{r table-display, echo=FALSE, eval=has_latex, out.width="45%"}
include_table("gear_cylinder_crosstab")
```


## Linear Models

The `o2f.lm()` method extracts coefficient tables from linear model objects.
Control which statistics appear using the `include` parameter.

```{r lm-example, eval=has_latex}
model <- lm(mpg ~ cyl + hp + wt, data = mtcars)

o2f(model,
    filename = "regression_table",
    caption = "Multiple Regression: MPG Predictors",
    include = c("estimate", "std.error", "p.value"),
    digits = 3,
    sub_dir = output_dir)
```

```{r lm-display, echo=FALSE, eval=has_latex, out.width="55%"}
include_table("regression_table")
```

Add confidence intervals (code only - full statistics):

```{r lm-ci-example, eval=FALSE}
o2f(model,
    filename = "regression_with_ci",
    include = c("estimate", "std.error", "conf.int", "p.value"),
    conf.level = 0.95,
    sub_dir = output_dir)
```


## Generalized Linear Models

The `o2f.glm()` method supports logistic, Poisson, and other GLM families.
Use `exponentiate = TRUE` to display odds ratios or rate ratios.

```{r glm-example, eval=has_latex}
logit_model <- glm(am ~ mpg + hp + wt,
                   data = mtcars,
                   family = binomial)

o2f(logit_model,
    filename = "logistic_odds_ratios",
    caption = "Logistic Regression: Odds Ratios",
    exponentiate = TRUE,
    include = c("estimate", "conf.int", "p.value"),
    sub_dir = output_dir)
```

```{r glm-display, echo=FALSE, eval=has_latex, out.width="55%"}
include_table("logistic_odds_ratios")
```


## ANOVA Tables

Both `aov` and `anova` objects are supported via `o2f.aov()` and `o2f.anova()`.

```{r aov-example, eval=has_latex}
aov_model <- aov(mpg ~ factor(cyl) + factor(gear), data = mtcars)

o2f(aov_model,
    filename = "anova_table",
    caption = "Two-way ANOVA: MPG by Cylinder and Gear",
    sub_dir = output_dir)
```

```{r aov-display, echo=FALSE, eval=has_latex, out.width="65%"}
include_table("anova_table")
```


## Hypothesis Tests

The `o2f.htest()` method handles output from t-tests, chi-square tests, and
other hypothesis testing functions.

```{r ttest-example, eval=has_latex}
ttest <- t.test(mpg ~ am, data = mtcars)

o2f(ttest,
    filename = "ttest_result",
    caption = "Two-sample t-test: MPG by Transmission",
    sub_dir = output_dir)
```

```{r ttest-display, echo=FALSE, eval=has_latex, out.width="70%"}
include_table("ttest_result")
```


## Regression Model Comparisons

The `o2f_regression()` function creates side-by-side model comparison tables
with significance stars.

```{r regression-comparison, eval=has_latex}
m1 <- lm(mpg ~ wt, data = mtcars)
m2 <- lm(mpg ~ wt + hp, data = mtcars)
m3 <- lm(mpg ~ wt + hp + cyl, data = mtcars)

o2f_regression(
  `(1)` = m1,
  `(2)` = m2,
  `(3)` = m3,
  stars = TRUE,
  digits = 3,
  filename = "model_progression",
  sub_dir = output_dir
)
```

```{r regression-comparison-display, echo=FALSE, eval=has_latex, out.width="65%"}
include_table("model_progression")
```


## RMS Package Models (v0.2.1)

zzobj2fig v0.2.1 adds full support for Frank Harrell's rms (Regression Modeling
Strategies) package, commonly used in biostatistics and clinical trials.

### rms Model Types

```{r rms-example, eval=FALSE}
library(rms)
dd <- datadist(mtcars)
options(datadist = "dd")

ols_model <- ols(mpg ~ rcs(wt, 4) + cyl, data = mtcars)
o2f(ols_model, filename = "ols_results")
o2f(ols_model, output = "anova", filename = "ols_chunk_tests")

lrm_model <- lrm(am ~ mpg + wt, data = mtcars)
o2f(lrm_model, exponentiate = TRUE, filename = "logistic_or")

library(survival)
dd <- datadist(lung)
options(datadist = "dd")
cph_model <- cph(Surv(time, status) ~ age + sex, data = lung)
o2f(cph_model, exponentiate = TRUE, filename = "cox_hr")
```

### rms Model Comparison

```{r rms-compare-example, eval=FALSE}
m1 <- ols(mpg ~ wt, data = mtcars)
m2 <- ols(mpg ~ wt + cyl, data = mtcars)
m3 <- ols(mpg ~ wt + cyl + hp, data = mtcars)

o2f_rms_compare(Simple = m1, Moderate = m2, Full = m3,
                filename = "model_comparison")
```

### Complete List of rms Methods

| Object Class | Method | Key Parameters |
|:-------------|:-------|:---------------|
| `ols` | `o2f.ols()` | `output` (coef/anova) |
| `lrm` | `o2f.lrm()` | `exponentiate`, `output` |
| `cph` | `o2f.cph()` | `exponentiate`, `output` |
| `orm` | `o2f.orm()` | `exponentiate`, `intercepts` |
| `Glm` | `o2f.Glm()` | `exponentiate`, `output` |
| `psm` | `o2f.psm()` | `output` |

**Note**: The `output = "anova"` option produces ANOVA-style chunk tests, which
is particularly useful for models with spline terms (e.g., `rcs()`).


## Additional Object Types via broom

zzobj2fig supports additional statistical object types through broom
integration. These methods use `broom::tidy()` to extract model information,
then format the results.

### Survival Analysis (survival package)

```{r survival-example, eval=FALSE}
library(survival)

cox_model <- coxph(Surv(time, status) ~ age + sex, data = lung)
o2f(cox_model, exponentiate = TRUE, filename = "cox_hr")

survreg_model <- survreg(Surv(time, status) ~ age + sex, data = lung)
o2f(survreg_model, filename = "parametric_survival")

km_fit <- survfit(Surv(time, status) ~ sex, data = lung)
o2f(km_fit, filename = "kaplan_meier")
```

### Mixed Effects Models (lme4, nlme packages)

```{r mixed-example, eval=FALSE}
library(lme4)

lmer_model <- lmer(mpg ~ cyl + (1 | gear), data = mtcars)
o2f(lmer_model, filename = "mixed_effects")

library(nlme)
lme_model <- lme(mpg ~ cyl, random = ~ 1 | gear, data = mtcars)
o2f(lme_model, filename = "nlme_model")
```

### Other Statistical Models

```{r other-models-example, eval=FALSE}
nls_model <- nls(mpg ~ a * exp(b * wt), data = mtcars,
                 start = list(a = 40, b = -0.5))
o2f(nls_model, filename = "nonlinear")

library(MASS)
polr_model <- polr(factor(gear) ~ mpg + hp, data = mtcars)
o2f(polr_model, filename = "ordinal_logistic")
```

### Exploratory Analysis

```{r exploratory-example, eval=FALSE}
pca_result <- prcomp(mtcars[, 1:5], scale. = TRUE)
o2f(pca_result, matrix = "rotation", filename = "pca_loadings")
o2f(pca_result, matrix = "summary", filename = "pca_summary")

km_result <- kmeans(mtcars[, 1:5], centers = 3)
o2f(km_result, matrix = "centers", filename = "cluster_centers")
```

### Complete List of broom-Based Methods

| Object Class | Package | Method | Key Parameters |
|:-------------|:--------|:-------|:---------------|
| `coxph` | survival | `o2f.coxph()` | `exponentiate`, `conf.int` |
| `survreg` | survival | `o2f.survreg()` | `digits` |
| `survfit` | survival | `o2f.survfit()` | `times` |
| `survdiff` | survival | `o2f.survdiff()` | `digits` |
| `Arima` | stats | `o2f.Arima()` | `digits` |
| `nls` | stats | `o2f.nls()` | `conf.int` |
| `polr` | MASS | `o2f.polr()` | `exponentiate`, `conf.int` |
| `multinom` | nnet | `o2f.multinom()` | `exponentiate`, `conf.int` |
| `prcomp` | stats | `o2f.prcomp()` | `matrix` (rotation/summary) |
| `kmeans` | stats | `o2f.kmeans()` | `matrix` (centers/summary) |
| `lmerMod` | lme4 | `o2f.lmerMod()` | `effects`, `conf.int` |
| `glmerMod` | lme4 | `o2f.glmerMod()` | `effects`, `exponentiate` |
| `lme` | nlme | `o2f.lme()` | `effects` |

**Note**: These methods require the `broom` package (or `broom.mixed` for
lme4/nlme models) and the respective model packages to be installed.


## Extensibility via o2f_tidy()

For any object type supported by `broom::tidy()` but without a dedicated
zzobj2fig method, use `o2f_tidy()`:

```{r o2f-tidy-example, eval=FALSE}
library(randomForest)
rf <- randomForest(Species ~ ., data = iris)
o2f_tidy(rf, filename = "rf_importance")

library(mgcv)
gam_model <- gam(mpg ~ s(wt) + s(hp), data = mtcars)
o2f_tidy(gam_model, filename = "gam_results")
```

This provides a fallback for the 50+ object types supported by broom that
don't have dedicated zzobj2fig methods.


# Theme Demonstrations

Each built-in theme provides distinct styling appropriate for different
publication contexts. Below are side-by-side comparisons using the same data.

```{r theme-data}
demo_data <- mtcars[1:6, c("mpg", "cyl", "hp", "wt")]
```

## Default Theme

The default styling uses blue row shading with booktabs rules.

```{r theme-default, eval=has_latex}
o2f(demo_data,
    filename = "theme_default",
    caption = "Default Theme",
    sub_dir = output_dir)
```

```{r theme-default-display, echo=FALSE, eval=has_latex, out.width="55%"}
include_table("theme_default")
```


## NEJM Theme

The New England Journal of Medicine theme provides clean, professional medical
journal styling with light yellow shading and sans-serif fonts.

```{r theme-nejm, eval=has_latex}
o2f(demo_data,
    filename = "theme_nejm",
    caption = "NEJM Theme",
    theme = "nejm",
    sub_dir = output_dir)
```

```{r theme-nejm-display, echo=FALSE, eval=has_latex, out.width="55%"}
include_table("theme_nejm")
```

NEJM theme characteristics:

- Light yellow alternating row colors
- Sans-serif font (Helvetica)
- Booktabs horizontal rules
- Compact column spacing (4pt)
- Footnotesize font


## APA Theme

The American Psychological Association theme follows APA 7th edition table
guidelines.

```{r theme-apa, eval=has_latex}
o2f(demo_data,
    filename = "theme_apa",
    caption = "APA Theme",
    theme = "apa",
    sub_dir = output_dir)
```

```{r theme-apa-display, echo=FALSE, eval=has_latex, out.width="55%"}
include_table("theme_apa")
```

APA theme characteristics:

- No row shading
- Times Roman font
- Letter paper with 1-inch margins
- Booktabs rules


## Nature Theme

The Nature journal theme uses Helvetica with subtle gray shading.

```{r theme-nature, eval=has_latex}
o2f(demo_data,
    filename = "theme_nature",
    caption = "Nature Theme",
    theme = "nature",
    sub_dir = output_dir)
```

```{r theme-nature-display, echo=FALSE, eval=has_latex, out.width="55%"}
include_table("theme_nature")
```

Nature theme characteristics:

- Light gray row shading (gray!8)
- Sans-serif font (Helvetica)
- Small font size
- 15mm margins


## Minimal Theme

The minimal theme provides clean, unobtrusive styling suitable for general
use.

```{r theme-minimal, eval=has_latex}
o2f(demo_data,
    filename = "theme_minimal",
    caption = "Minimal Theme",
    theme = "minimal",
    sub_dir = output_dir)
```

```{r theme-minimal-display, echo=FALSE, eval=has_latex, out.width="55%"}
include_table("theme_minimal")
```


# Custom Themes

Create custom themes for specific requirements using `o2f_theme()`.

```{r custom-theme-create, eval=has_latex}
presentation_theme <- o2f_theme(
  name = "presentation",
  scolor = "blue!15",
  header_bold = TRUE,
  font_size = "large",
  striped = TRUE,
  extra_packages = list(
    geometry(margin = "10mm"),
    "\\renewcommand{\\arraystretch}{1.3}"
  )
)

o2f(demo_data,
    filename = "theme_presentation",
    caption = "Custom Presentation Theme",
    theme = presentation_theme,
    sub_dir = output_dir)
```

```{r custom-theme-display, echo=FALSE, eval=has_latex, out.width="70%"}
include_table("theme_presentation")
```


# Global Theme Setting

Set a theme globally for consistent styling across multiple tables in an
analysis.

```{r global-theme-workflow, eval=FALSE}
o2f_theme_set("nejm")

o2f(mtcars[1:5, 1:4], filename = "table1", sub_dir = "output")
o2f(iris[1:5, ], filename = "table2", sub_dir = "output")

model <- lm(mpg ~ wt + hp, data = mtcars)
o2f(model, filename = "table3", sub_dir = "output")

o2f(mtcars[1:5, 1:4],
    filename = "table4_apa",
    theme = "apa",
    sub_dir = "output")

o2f_theme_set(NULL)
```


# Summary

## Object Type Support (Base Methods)

| Object Class | Method | Key Parameters |
|:-------------|:-------|:---------------|
| `data.frame` | `o2f.data.frame()` | Standard o2f parameters |
| `matrix` | `o2f.matrix()` | `rownames` |
| `table` | `o2f.table()` | Standard o2f parameters |
| `lm` | `o2f.lm()` | `digits`, `include`, `conf.level` |
| `glm` | `o2f.glm()` | `digits`, `include`, `exponentiate`, `conf.level` |
| `aov` | `o2f.aov()` | `digits` |
| `anova` | `o2f.anova()` | `digits` |
| `htest` | `o2f.htest()` | `digits` |

## Object Type Support (rms package)

| Object Class | Method | Key Parameters |
|:-------------|:-------|:---------------|
| `ols` | `o2f.ols()` | `output` (coef/anova) |
| `lrm` | `o2f.lrm()` | `exponentiate`, `output` |
| `cph` | `o2f.cph()` | `exponentiate`, `output` |
| `orm` | `o2f.orm()` | `exponentiate`, `intercepts`, `output` |
| `Glm` | `o2f.Glm()` | `exponentiate`, `output` |
| `psm` | `o2f.psm()` | `output` |

## Object Type Support (via broom)

| Object Class | Package | Key Parameters |
|:-------------|:--------|:---------------|
| `coxph`, `survreg`, `survfit`, `survdiff` | survival | `exponentiate` |
| `nls` | stats | `conf.int` |
| `polr`, `multinom` | MASS, nnet | `exponentiate` |
| `prcomp`, `kmeans` | stats | `matrix` |
| `lmerMod`, `glmerMod` | lme4 | `effects` |
| `lme` | nlme | `effects` |
| `Arima` | stats | `digits` |
| Any broom object | `o2f_tidy()` | `tidy_args`, `glance` |

## Theme Comparison

| Theme | Row Shading | Font | Use Case |
|:------|:------------|:-----|:---------|
| Default | blue!10 | Serif | General purpose |
| NEJM | yellow!8 | Helvetica | Medical journals |
| APA | None | Times | Psychology/social sciences |
| Nature | gray!8 | Helvetica | Scientific journals |
| Minimal | gray!5 | Serif | Clean, simple tables |

## Theme Functions

| Function | Purpose |
|:---------|:--------|
| `o2f_theme()` | Create custom theme |
| `o2f_theme_set()` | Set global theme |
| `o2f_theme_get()` | Get current global theme |
| `o2f_theme_nejm()` | NEJM preset |
| `o2f_theme_apa()` | APA preset |
| `o2f_theme_nature()` | Nature preset |
| `o2f_theme_minimal()` | Minimal preset |
| `o2f_list_themes()` | List available themes |
